Hypertension and microbiome

### 2019-01-23 articletwo.Rmd
Minor

@@

```{r Write characteristics}
tbl1head <- "Characteristics of the study sample."
tbl1foot <- "Continuous variables are presented as mean (standard deviation). Categorical variables reported as absolute and relative frequencies. BP indicates blood pressure, BMI body mass index, RAS renin-angiotensin system."
writetable(doc, tbl1, number = 1, tbl1head, tbl1foot)
```

@@

ggsave(file = "cache/alpha-beta.png", plot=g4, height=fig.height, width=11)
writeimage(doc, "cache/alpha-beta.png", "Alpha-beta", "footer")
```

@@

pcaplot <- ggplot(cbind(meta(pseq.species), pca.prcomp$x[,1:2]), aes(x=PC1, y=PC2)) +
    geom_point(aes(color = HYPERTENSION), size = 0.5) +
    scale_color_manual(values=c("red", "blue"),
                       name= "Hypertension",
                       breaks=c("1", "0"),
                       labels=c("Yes", "No")) +
    xlab(pca.labels[1]) +
    ylab(pca.labels[2]) +
    coord_fixed() +
    scale_x_continuous(breaks=seq(-100, 100, 20)) +
    scale_y_continuous(breaks=seq(-30, 30, 10)) +
    theme_classic() +
    theme(text = element_text(size=20),
          legend.justification=c(1,0),
          legend.position=c(1.03, 0.75),
          legend.title=element_text(size = 20),
          legend.direction="vertical",
          legend.text=element_text(size = floor(0.75*20)),
          legend.box.margin = margin(1, 1, 1, 1, "mm"),
          legend.box.background = element_rect(colour = "black"))

@@

writeimage(doc, "cache/pca.png", "Alpha-beta", "footer", width = 7)
```

@@

writeggplot(doc, maaslinplot, "maaslin", "footer", width = 7, height = 3)
```

@@

```{r PCA loadings}

pca.rotation.abs <- abs(pca.prcomp$rotation)
pca.loading <- sweep(pca.rotation.abs, 2, colSums(pca.rotation.abs), "/") %>%
    as.data.frame %>%
    tibble::rownames_to_column(var = "term")

@@

pca.loading %>% pull(PC3) %>% head(2700) %>% sum

ggplot(data=pca.pc3, aes(x=reorder(term, -PC3), y=PC3)) +
    geom_bar(stat="identity") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

dev.off()

>>>>

Session info

```{r session info}
sessionInfo()
```

@@

```{r Write characteristics}
tbl1head <- "Characteristics of the study sample."
tbl1foot <- "Continuous variables are presented as mean (standard deviation). Exercise at free-timeis reported as not much (1), walk etc. (2), regular three hours or more or competitive sports (3). Categorical variables reported as absolute and relative frequencies. BP indicates blood pressure, BMI body mass index, RAS renin-angiotensin system."
writetable(doc, tbl1, number = 1, tbl1head, tbl1foot)
```

@@

ggsave(file = "cache/alpha-beta.png", plot=g4, height=fig.height, width=11)

@@

pcaplot <- ggplot(cbind(meta(pseq.species), pca.prcomp$x[,1:2]), aes(x=PC1, y=PC2)) +
    geom_point(aes(color = HYPERTENSION), size = 0.5) +
    scale_color_manual(values=c("blue", "red"),
                       name= "Hypertension",
                       breaks=c("1", "0"),
                       labels=c("Yes", "No")) +
    xlab(pca.labels[1]) +
    ylab(pca.labels[2]) +
    coord_fixed() +
    scale_x_continuous(breaks=seq(-100, 100, 20)) +
    scale_y_continuous(breaks=seq(-30, 30, 10)) +
    theme_classic() +
    theme(text = element_text(size=20),
          legend.justification=c(1,0),
          legend.position=c(1.03, 0.75),
          legend.title=element_text(size = 20),
          legend.direction="vertical",
          legend.text=element_text(size = floor(0.75*20)),
          legend.box.margin = margin(1, 1, 1, 1, "mm"),
          legend.box.background = element_rect(colour = "black"))

@@

writeimage(doc, 2, "cache/pca.png", "Alpha-beta", "footer", width = 7)
```

```{r Write figure 2}
fig2head <- "Tenth of bacterial variation can be explained by two first principal axes"
fig2foot <- sprintf("Principal component analysis for bacterial abundances at species level. Centered log ratio transform is performed for the compositional data. First two axes explain %.1f%% of the variation. Research participants with hypertension are denoted by red dot.", sum(pca.percentage[1:2]))
writeimage(doc, 2, "cache/pca.png", fig2head, fig2foot)
```

## Direct associations between genus level and blood pressure variables (figure 3)

@@

ggsave(file = "cache/maaslin.png", plot = maaslinplot, height = 3, width = 7, dpi = 300)
```

```{r Write figure 3}
fig3head <- ""
fig3foot <- ""
writeimage(doc, 3, "cache/maaslin.png", fig3head, fig3foot)
```

## Figure 4

@@

## Loadings for principal component axis 3 (Figure 5)

```{r Main contributors of PC3}
pca.rotation.abs <- abs(pca.prcomp$rotation)
pca.loading <- sweep(pca.rotation.abs, 2, colSums(pca.rotation.abs), "/") %>%
    as.data.frame %>%
    tibble::rownames_to_column(var = "term")

pca.prcomp$rotation %>% length

pca.loading %>% pull(PC3) %>% sum

@@

pca.pc3.gg <- ggplot(data=pca.pc3, aes(x=reorder(term, -PC3), y=PC3)) +
    geom_bar(stat="identity") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggsave(file = "cache/pca.pc3.png", plot = pca.pc3.gg, height = 3, width = 7, dpi = 300)
```

```{r Write figure 4}
fig4head <- ""
fig4foot <- ""
writeimage(doc, 4, "cache/pca.pc3.png", fig4head, fig4foot)
```

### 2019-01-23 rrbiome.Rmd
Minor

 ```{r adonis calculation, cache = FALSE}
if (!file.exists("rds/adonis.species.rds")) {

    adonis.species <- calculateadonis(dset = meta(pseq.species),
                    matrix = bray.dist.m.species,
                        covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
                    npermutations = 1)

    saveRDS(adonis.species, file = "rds/adonis.species.rds")

} else {
     adonis.species  <- readRDS("rds/adonis.species.rds")
}
```

>>>>

 ```{r adonis calculation, cache = FALSE}
if (!file.exists("rds/adonis.species.rds")) {
    adonis.species <- calculateadonis(dset = meta(pseq.species),
                                      matrix = bray.dist.m.species,
                                      covariates = c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"),
                                      npermutations = 99)
    saveRDS(adonis.species, file = "rds/adonis.species.rds")
    } else {
     adonis.species  <- readRDS("rds/adonis.species.rds")
}
```
